### 2013-06-25 MIMOSA/mimosafigures/mimosafigures-test/manuscriptFigures.Rmd

Load the data and fit the MIMOSA model
===============

```{r }
source("utility.R")
```{r echo=FALSE}
setup()
```

### Read and merge the hvtn data.

```{r cache=TRUE}
hvtndata<-readHVTNData()
hvtndata<-processHVTNData(hvtndata)
hvtndata<-constructEset(hvtndata)
```

### Read the fluidigm data.

```{r cache=FALSE}
fluidigm<-readFluidigmData()
```

Process it.

```{r cache=FALSE}
fluidigm<-processFluidigmData()
```

### Fit the MIMOSA model to the HVTN data all cytokines and ENV-1-PTEG using EM and MCMC implementations.

```{r }
if(!file.exists("hvtnEMfit.rds")){
  hvtn.result.EM<-MIMOSA(NSUB+CYTNUM~ASSAYID+VISITNO+PTID+RX_CODE+RefTreat|TCELLSUB+ANTIGEN+CYTOKINE,hvtndata,ref=RefTreat%in%"Reference"&ANTIGEN%in%"ENV-1-PTEG",subset=RefTreat%in%"Treatment"&ANTIGEN%in%"ENV-1-PTEG",method="EM",run.parallel=TRUE)
  saveRDS(hvtn.result.EM,file="hvtnEMfit.rds")
}else{
  hvtn.result.EM<-readRDS("hvtnEMfit.rds")
}
if(!file.exists("hvtnMCMCfit.rds")){
  hvtn.result.mcmc<-MIMOSA(NSUB+CYTNUM~ASSAYID+VISITNO+PTID+RX_CODE+RefTreat|TCELLSUB+ANTIGEN+CYTOKINE,hvtndata,ref=RefTreat%in%"Reference"&ANTIGEN%in%"ENV-1-PTEG",subset=RefTreat%in%"Treatment"&ANTIGEN%in%"ENV-1-PTEG",method="mcmc",run.parallel=TRUE,pXi=0.1,getP=TRUE)
  saveRDS(hvtn.result.mcmc,file="hvtnMCMCfit.rds")
}else{
  hvtn.result.mcmc<-readRDS("hvtnMCMCfit.rds")
}
```

### Summarize the results. Do the likelihood ratio test, Fisher's test, and rank by log fold change for comparison.

```{r cache=TRUE,message=FALSE,warning=FALSE}
hvtn.summary.EM<-summarizeMIMOSAFit(hvtn.result.EM)
hvtn.summary.EM<-subset(hvtn.summary.EM,Method%in%"EM")
hvtn.summary.mcmc<-summarizeMIMOSAFit(hvtn.result.mcmc)
hvtn.summary.mcmc<-FisherFromSummary(hvtn.summary.mcmc)
hvtn.summary.mcmc<-LRT(hvtn.summary.mcmc)
hvtn.summary.mcmc<-FoldChangeRank(hvtn.summary.mcmc)
```

### Compute ROC and FDR curves and calculate the AUC for all the results.

```{r cache=TRUE}
ROC<-ComputeROCs(hvtn.summary.mcmc,hvtn.summary.EM)
FDR<-ComputeFDRs(hvtn.summary.mcmc,hvtn.summary.EM)
AUC<-ComputeAUCs(ROC)
```

### FDR and ROC plots of the ICS data.
--------
We see that MIMOSA performs as well or better than competing methods. The MCMC routine performs better than the EM routine for some data.

```{r cache=TRUE,fig.width=10,fig.height=8,fig.cap='The TNF expressing cytokine specific cells have high background as evidenced by the poor performance of all methods fit to TNF+ subsets of cells, with the exception of the triple-positive cell subset. As the number of simultaneously expressed cytokines being examined increases, the background decreases.'}
ROC$Method<-relevel(ROC$Method,"MIMOSA (EM)")
levs <-
  c("IFN-IL2-TNF+", "IFN-IL2+TNF-", "IFN-IL2+TNF+", "IFN+IL2-TNF-", 
    "IFN+IL2-TNF+", "IFN+IL2+TNF-", "IFN+IL2+TNF+", "IFNg+", "IFNg+IL2+", 
    "IFNg+TNF+", "IL2+", "IL2-IFNg+", "IL2+ OR IFNg+", "IL2+IFNg-", 
    "IL2+IFNg+", "IL2+TNF+", "TNF+")
ROC$CYTOKINE<-factor(ROC$CYTOKINE,labels=levs)
AUC$CYTOKINE<-factor(AUC$CYTOKINE,labels=levs)
rocplots<-ggplot(subset(ROC,TCELLSUB%in%"cd3+/cd4+"))+geom_line(aes(x=FPR,y=TPR,color=Method))+theme_bw()+facet_wrap(~CYTOKINE)+geom_text(aes(x=x,y=y,label=sprintf("AUC=%s",signif(AUC,2)),col=Method),size=3,data=AUC,show_guide=FALSE)+theme(axis.text.x=element_text(angle=90,hjust=1))
```{r fig.width=10,fig.height=8,fig.cap='The FDR is well controlled for cytokine subsets with low background. The TNF expressing subsets have high background and consequently positive assay results in the "Visit 2" samples, which are labelled as "true negatives" in this scheme. The result is observed as a large observed false discovery rate for all methods.'}
FDR$Method<-relevel(FDR$Method,"MIMOSA (EM)")
levs<-levels(FDR$CYTOKINE)
levs <-
  c("IFN-IL2-TNF+", "IFN-IL2+TNF-", "IFN-IL2+TNF+", "IFN+IL2-TNF-", 
    "IFN+IL2-TNF+", "IFN+IL2+TNF-", "IFN+IL2+TNF+", "IFNg+", "IFNg+IL2+", 
    "IFNg+TNF+", "IL2+", "IL2-IFNg+", "IL2+ OR IFNg+", "IL2+IFNg-", 
    "IL2+IFNg+", "IL2+TNF+", "TNF+")
FDR$CYTOKINE<-factor(FDR$CYTOKINE,labels=levs)
fdrplots<-ggplot(subset(FDR,TCELLSUB%in%c("cd3+/cd4+")))+geom_line(aes(y=true.fdr,x=fdr,color=Method))+geom_abline(1,lty=3)+theme_bw()+scale_x_continuous(name="Nominal FDR")+scale_y_continuous(name="Observed FDR")+facet_wrap(~CYTOKINE)+theme(axis.text.x=element_text(angle=90,hjust=1))
rocplots
fdrplots
```

### Figure 1 and S1.

```{r echo=FALSE,eval=FALSE}
suppressPackageStartupMessages(library(gridExtra))
f1levs<-levs[c(8,11,13,17,15,16,10,7)]
sflevs<-setdiff(levs,f1levs)
AUC$y[AUC$Method%in%"Fisher"]<-0.4
AUC$y[AUC$Method%in%"MIMOSA (EM)"]<-0.5
AUC$y[AUC$Method%in%"MIMOSA (MCMC)"]<-0.3

fnames<-c("revised_2.pdf","revised_1.pdf","revised_15.pdf","revised_3.pdf","revised_5.pdf","revised_6.pdf","revised_7.pdf","revised_8.pdf")
for(i in seq_along(f1levs)){
  fi.1<-ggplot(subset(ROC,TCELLSUB%in%"cd3+/cd4+"&CYTOKINE%in%f1levs[i]))+geom_line(aes(x=FPR,y=TPR,color=Method))+theme_bw()+facet_wrap(~CYTOKINE)+geom_text(aes(x=x,y=y,label=sprintf("AUC=%s",signif(AUC,2)),col=factor(Method)),data=subset(AUC,CYTOKINE%in%f1levs[i]),show_guide=FALSE)+theme(axis.text.x=element_text(angle=90,hjust=1)) 
  fi.2<-ggplot(subset(FDR,TCELLSUB%in%c("cd3+/cd4+")&CYTOKINE%in%f1levs[i]))+geom_line(aes(y=true.fdr,x=fdr,color=Method))+geom_abline(1,lty=3)+theme_bw()+scale_x_continuous(name="Nominal FDR")+scale_y_continuous(name="Observed FDR")+facet_wrap(~CYTOKINE)+theme(axis.text.x=element_text(angle=90,hjust=1))
  pdf(file=sprintf("/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/%s",fnames[i]),width=10,height=3)
  grid.arrange(fi.1,fi.2,ncol=2,nrow=1)
  dev.off()
  grid.arrange(fi.1,fi.2,ncol=2,nrow=1)
}
```

### Process the fluidigm data to construct an expression set for MIMOSA.
---------

```{r warning=FALSE,cache=TRUE}
fl.bystim<-lapply(fluidigm,exprs)
cd.bystim<-lapply(fluidigm,cData)
fl<-lapply(names(fl.bystim),function(n){
  m<-melt(cbind(fl.bystim[[n]],cd.bystim[[n]]),id=c("Stim.Agent","Stim.Condition","Time.of.Stim","Sero.Status","Patient.ID","Chip.Number","Well","Number.of.Cells"))
  ddply(m,.(Stim.Condition,Patient.ID,variable),summarize,N=length(value),n=sum(value>0))
})
fl<-lapply(fl,function(x)setnames(x,"variable","Gene"))
Esets.fluidigm<-lapply(fl,function(x)ConstructMIMOSAExpressionSet(x,reference=Stim.Condition%in%"Unstim",measure.columns=c("N","n"),other.annotations=c("Gene","Patient.ID","Stim.Condition"),default.cast.formula=component~Patient.ID+Gene,.variables=.(Patient.ID,Gene),featureCols=1,ref.append.replace="_REF"))
nms<-names(fl.bystim)
Esets.fluidigm<-lapply(1:length(Esets.fluidigm),function(x){pData(Esets.fluidigm[[x]])<-cbind(pData(Esets.fluidigm[[x]]),Stim=nms[x]);Esets.fluidigm[[x]]})
Eset.combined<-combine(Esets.fluidigm[[1]],Esets.fluidigm[[2]],Esets.fluidigm[[3]],Esets.fluidigm[[4]])
```

### Fit MIMOSA by gene, grouping by stimulation and for all samples together.
----

```{r fluidigm.fit, cache=TRUE}
if(!file.exists("fluidigm.fits.by.stim.allgenes.rds")){
  fluidigm.fits.by.stim.allgenes<-lapply(list(Eset.combined),function(x)MIMOSA(N+n~Stim+Patient.ID|Gene,x,ref=RefTreat%in%"Reference",subset=RefTreat%in%"Treatment",method="mcmc",EXPRATE=1e-4,iter=105000,burn=5000,pXi=1,alternative="not equal",run.parallel=TRUE,getP=TRUE))
  fluidigm.fits.by.stim.bygene<-mclapply(Esets.fluidigm,function(x)MIMOSA(N+n~Stim+Patient.ID|Gene,x,ref=RefTreat%in%"Reference",subset=RefTreat%in%"Treatment",method="mcmc",EXPRATE=1e-4,iter=105000,burn=50000,pXi=1,alternative="not equal",getP=TRUE,run.parallel=TRUE))
  saveRDS(fluidigm.fits.by.stim.allgenes,file="fluidigm.fits.by.stim.allgenes.rds")
  saveRDS(fluidigm.fits.by.stim.bygene,file="fluidigm.fits.by.stim.bygene.rds")
}else{
  fluidigm.fits.by.stim.allgenes<-readRDS("fluidigm.fits.by.stim.allgenes.rds")
  fluidigm.fits.by.stim.bygene<-readRDS("fluidigm.fits.by.stim.bygene.rds")
}
```

### Heatmaps for Fluidigm data fit to all stimulations simultaneously.
----
-Red and orange are CMV (pp65). Yellow and green are HIV (gag and nef).
-When fitting the data to all samples simultaneously, we still observe clustering of samples by stimulation (i.e. CMV stimulated samples cluster together, and HIV stimulated samples cluster together).

```{r heatmap.probs.bygene,cache=TRUE}
foo<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)x@z[,2]))
props<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)do.call(c,lapply(x@result@p,function(x)-diff(x[,2])))))
S<-sign(props)
S[S==0]<-1
colnames(S)<-pData(fluidigm.fits.by.stim.allgenes[[1]][[1]])[,c("Patient.ID")]
stim<-pData(fluidigm.fits.by.stim.allgenes[[1]][[1]])$Stim
rownames(S)<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)as.character(unique(pData(x)$Gene))))
#heatmap.2((S*(foo))[apply(foo,1,function(x)sum(x>0.6)>=1),],col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="signed Posterior Probabilities\n(fit by gene, all stimulations)")
foo2<-apply(foo,2,function(x)MIMOSA:::fdr(cbind(1-x,x)))
heatmap.2((S*(foo))[apply(foo2,1,function(x)sum(x<0.1)>=1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.75,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
```

### Figure 2 B, (fit of all samples simultaneously)

```{r eval=FALSE,echo=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/FluidigmCombinedFit.pdf")
foo2<-apply(foo,2,function(x)MIMOSA:::fdr(cbind(1-x,x)))
heatmap.2((S*(foo))[apply(foo2,1,function(x)sum(x<0.1)>=1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.75,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
dev.off()
```

Heatmap of posterior differences in proportions from simultaneous fit (not shown).
----
-Red and orange are CMV (pp65). Yellow and green are HIV (gag and nef).
-When fitting the data to all samples simultaneously, we still observe clustering of samples by stimulation (i.e. CMV stimulated samples cluster together, and HIV stimulated samples cluster together).

```{r heatmap.props.bygene,cache=TRUE,include=FALSE}
colnames(props)<-colnames(S)
rownames(props)<-rownames(S)
heatmap.2(S*sqrt(abs(props)),col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="Posterior Differences in Proportions\n(fit by gene all stimulations)")
```

### Heatmaps for Fluidigm data fit per stimulation.
---
-Red and orange are CMV (pp65). Yellow and green are HIV (gag and nef).

```{r heatmap.probs.bystim.bygene,cache=TRUE}
foo<-do.call(cbind,lapply(fluidigm.fits.by.stim.bygene,function(x)do.call(rbind,lapply(x,function(x)x@z[,2]))))
props<-do.call(cbind,lapply(fluidigm.fits.by.stim.bygene,function(x)do.call(rbind,lapply(x,function(x)do.call(c,lapply(x@result@p,function(x)-diff(x[,2])))))))
S<-sign(props)
colnames(S)<-do.call(c,lapply(fluidigm.fits.by.stim.bygene,function(x)as.character(pData(x[[1]])[,c("Patient.ID")])))
S[S==0]<-1
stim<-as.numeric(factor(do.call(c,lapply(fluidigm.fits.by.stim.bygene,function(x)as.character(pData(x[[1]])$Stim)))))
colnames(S)<-do.call(c,lapply(fluidigm.fits.by.stim.bygene,function(x)as.character(pData(x[[1]])$Patient.ID)))
rownames(S)<-do.call(cbind,lapply(fluidigm.fits.by.stim.bygene[[1]],function(x)as.character(unique(pData(x)$Gene))))
#h<-heatmap.2((S*foo)[apply(foo,1,function(x)sum(x>.6)>=1),],col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],main="Signed Posterior Probabilities\n(fit by gene and stimulation).")
foo2<-apply(foo,2,function(x)MIMOSA:::fdr(cbind(1-x,x)))
h<-heatmap.2((S*(foo))[apply(foo2,1,function(x)sum(x<0.1)>=1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.75,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
hmaporder<-h$colInd
geneinds<-apply(foo,1,function(x)sum(x>.6)>=1)
```

### Figure 2 A (fit per stimulation)
Fit MIMOSA per stimulation.

```{r eval=FALSE,echo=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Fluidigm_plot1.pdf")
foo2<-apply(foo,2,function(x)MIMOSA:::fdr(cbind(1-x,x)))
heatmap.2((S*(foo))[apply(foo2,1,function(x)sum(x<0.1)>=1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.75,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
dev.off()
```

Heatmap of posterior differences in proportions for fit of samples by stimulation and by gene (not shown).
---
-Red and orange are CMV (pp65). Yellow and green are HIV (gag and nef).

```{r heatmap.props.bystim.bygene,cache=TRUE,include=FALSE}
colnames(props)<-colnames(S)
rownames(props)<-rownames(S)
heatmap.2(sqrt(abs(props))*S,col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="Posterior differences in proportions\n(fit by gene and stimulations)")
```

### Fisher's exact test and heatmaps of -log10(q-values) from Fisher's test.
---
There is little clustering by stimulation evident from testing for differential expression with Fisher's exact test.

```{r cache=TRUE}
fisher<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x){
  p<-vector('numeric',nrow(x@result@n.stim))
  for(i in 1:nrow(x@result@n.stim)){
    p[i]<-fisher.test(as.matrix(rbind(x@result@n.stim[i,],x@result@n.unstim[i,])),alternative="two.sided")$p.value
  }
  p
}))
adj.fisher<-matrix(p.adjust(fisher,"fdr"),nrow=96)
S<-sign(lapply(fluidigm.fits.by.stim.allgenes,function(x)do.call(rbind,lapply(x,function(x)(prop.table(as.matrix(x@result@n.stim),1)-prop.table(as.matrix(x@result@n.unstim),1))[,2])))[[1]])
S[S==0]<-1
colnames(fisher)<-gsub("_B3GAT1_Treatment","",colnames(S))
rownames(fisher)<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x)as.character(unique(pData(x)$Gene))))
stim<-pData(fluidigm.fits.by.stim.allgenes[[1]][[1]])$Stim
#heatmap.2(fisher*S,col=colorpanel(n=10,high="yellow",low="blue",mid="white"),margins=c(6,5),cexRow=0.5,hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,main="-log10(p-values)\nfrom Fisher's Exact Test")
colnames(adj.fisher)<-colnames(fisher)
rownames(adj.fisher)<-rownames(fisher)
heatmap.2((((adj.fisher))*S)[apply(adj.fisher,1,function(x)sum(x<0.1)>1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
```

### Figure 2 D (q-values from Fisher's test)

```{r eval=FALSE,echo=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/FluidigmFisher.pdf")
colnames(adj.fisher)<-colnames(fisher)
rownames(adj.fisher)<-rownames(fisher)
heatmap.2((((adj.fisher))*S)[apply(adj.fisher,1,function(x)sum(x<0.1)>1),],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),hclustfun=function(x)hclust(x,method="complete"),trace="none",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)],Colv=TRUE,denscol="black")
dev.off()

```{r eval=TRUE,echo=FALSE}
emprop<-do.call(rbind,lapply(fluidigm.fits.by.stim.allgenes[[1]],function(x){
  p<-vector('numeric',nrow(x@result@n.stim))
  p<-(prop.table(as.matrix((x@result@n.stim)),1)-prop.table(as.matrix(x@result@n.unstim),1))[,2]
}))
rownames(emprop)<-rownames(fisher)
colnames(emprop)<-colnames(fisher)

```{r echo=FALSE,warning=FALSE}
heatmap.2(emprop[geneinds,hmaporder],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),Colv=NULL,denscol="black",tracecol="black",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)[hmaporder]])
```

### Figure 2C empirical proportions.

```{r eval=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/FluidigmEmpiricalProps.pdf")
heatmap.2(emprop[geneinds,hmaporder],col=colorpanel(n=100,high="yellow",low="blue",mid="white"),margins=c(6,5),Colv=NULL,denscol="black",tracecol="black",symbreaks=TRUE,ColSideColors=c("red","orange","yellow","green")[as.numeric(stim)[hmaporder]])
dev.off()
```

### Multivariate MIMOSA fit to fluidigm data, looking at interactions of genes.
---------
There is a difference when comparing stimulations and looking at the multivariate vector of expression for GZMK and CCR7 (i.e [+/+,+/-,-/+,--]), but no difference is detected when we look at expression of GZMK | CCR7, since one cell subset increases while the other decreases upon stimultion.

```{r cache=TRUE}
fl.comb<-combine(fluidigm[[1]],fluidigm[[2]],fluidigm[[3]],fluidigm[[4]])
melted<-melt(fl.comb)
require(glmnet)
require(Matrix)
melted$Et.bin<-as.numeric(as.logical(melted$Et))
#setnames(melted,"__wellKey","wellKey")
melted<-within(melted,group<-Patient.ID:Stim.Agent)
melted<-cast(melt(melted,id=c("group","Stim.Condition","primerid","wellKey"),measure="Et.bin"),group+Stim.Condition+wellKey~primerid)

# interact.scan<-combn(colnames(melted)[4:99],2,function(x){
#   g1<-x[1];g2<-x[2]
#   gg<-paste(g1,g2,sep=":")
#   bar<-data.frame(melted[,c("group","Stim.Condition",g1,g2)])
#   bar$foo<-factor(bar[,c(g1)]):factor(bar[,c(g2)])
#   setnames(bar,"foo",gg)
#   bar[,5]<-factor(bar[,5],levels=c("0:0","0:1","1:0","1:1"))
#   C<-cast(melt(bar,id=c("group","Stim.Condition"),measure=c(gg)),group+Stim.Condition~value,fun.aggregate=function(x)length(x))
#   expected<-c("0:0","0:1","1:0","1:1")
#   unused<-setdiff(expected,colnames(C))
#   newframe<-NULL
#   for(i in seq_along(expected)){
#     if(expected[i]%in%unused){
#       newframe<-cbind(newframe,rep(0,nrow(C)))
#     }else{
#       newframe<-cbind(newframe,C[,expected[i]])
#     }
#   }
#   colnames(newframe)<-expected
#   C<-cbind(C[,1:2],newframe)
#   CC<-list(n.unstim=subset(C,Stim.Condition%in%"Unstim")[,-c(1:2)],n.stim=subset(C,Stim.Condition%in%"Stim")[,-c(1:2)])
#   foo<-lapply(CC,function(x)prop.table(as.matrix(x),1))
#   psum<-mean(rowSums(foo[[1]][,2:4])-rowSums(foo[[2]][,2:4]))
#   pij<-colMeans(foo[[1]][,2:4]-foo[[2]][,2:4])
#   res<-c(psum,pij)
#   res
# })
#cache("interact.scan")

#genes<-combn(colnames(melted)[6:101],2)[,which(abs(interact.scan[1,])<0.01&abs(interact.scan[4,])>0.3)]
#interact.scan[,which(abs(interact.scan[1,])<0.01&abs(interact.scan[4,])>0.3)]
#ggg<-genes[,6]
#g1<-ggg[1];g2<-ggg[2] #CCR7 and GZMK
g1<-"CCR7";g2<-"GZMK"
gg<-paste(g1,g2,sep=":")
bar<-data.frame(melted[,c("group","Stim.Condition",g1,g2)])
bar$foo<-factor(bar[,c(g1)]):factor(bar[,c(g2)])
setnames(bar,"foo",gg)
C<-cast(melt(bar,id=c("group","Stim.Condition"),measure=c(gg)),group+Stim.Condition~value,fun.aggregate=function(x)length(x))
CC<-list(n.unstim=subset(C,Stim.Condition%in%"Unstim")[,-c(1:2)],n.stim=subset(C,Stim.Condition%in%"Stim")[,-c(1:2)])

toplot<-data.frame(rbind(data.frame(CC$n.stim),data.frame(CC$n.unstim)),stim=c(rep("Stim",nrow(CC$n.stim)),rep("Unstim",nrow(CC$n.unstim))))
toplot$id<-rep(unlist(lapply(strsplit(as.character(subset(C,Stim.Condition%in%"Unstim")[,1]),":"),function(x)x[1]),use.names=FALSE),2)
colnames(toplot)[1:4]<-gsub("X","",paste(g1,g2,colnames(toplot)[1:4],sep=":"))
toplot<-melt(toplot,id=c("id","stim"))
p1<-ggplot(toplot)+geom_tile(aes(x=variable,y=id,fill=log2(value)),interpolate=FALSE)+theme_bw()+theme(axis.text.x=element_text(angle=90))+facet_wrap(~stim)+scale_fill_continuous(na.value="gray")

DD<-CC
DD[[1]]<-cbind(CC[[1]][,1],rowSums(CC[[1]][,2:4]))
DD[[2]]<-cbind(CC[[2]][,1],rowSums(CC[[2]][,2:4]))
toplot2<-data.frame(rbind(data.frame(DD$n.stim),data.frame(DD$n.unstim)),stim=c(rep("Stim",nrow(DD$n.stim)),rep("Unstim",nrow(DD$n.unstim))))
toplot2$id<-rep(unlist(lapply(strsplit(as.character(subset(C,Stim.Condition%in%"Unstim")[,1]),":"),function(x)x[1]),use.names=FALSE),2)
colnames(toplot2)[1:2]<-gsub("X","",paste(g1,g2,colnames(toplot2)[1:2],sep=":"))
toplot2<-melt(toplot2,id=c("id","stim"))
p2<-ggplot(toplot2)+geom_tile(aes(x=variable,y=id,fill=log2(value)),interpolate=FALSE)+theme_bw()+theme(axis.text.x=element_text(angle=90))+facet_wrap(~stim)+scale_fill_continuous(na.value="gray")

```{r fitmulti,echo=FALSE,message=FALSE,warning=FALSE,error=FALSE,results='hide'}
flres<-.fitMCMC(data=CC,inits=MIMOSA:::MDMix(CC,initonly=TRUE),EXPRATE=1e-2,pXi=1,alternative="not equal")
flsum<-.fitMCMC(data=DD,inits=MIMOSA:::MDMix(DD,initonly=TRUE),EXPRATE=1e-2,pXi=1,alternative="not equal")
sapply(list.files(pattern=c("\\.dat")),unlink)
```

### Figure 4. Heatmaps of counts for all combinations of CCR7 and GZMK vs marginal counts.
Samples in red show no difference by MIMOSA.

```{r eval=TRUE,echo=FALSE}
library(gridExtra)
missed<-!MIMOSA:::fdr(flres$z)<0.05
toplot$variable<-factor(toplot$variable,labels=c("CCR7-GZMK-","CCR7-GZMK+","CCR7+GZMK-","CCR7+GZMK+"))
toplot2$variable<-factor(toplot2$variable,labels=c("CCR7-GZMK-","CCR7+ or GZMK+"))
p1<-ggplot(toplot)+geom_tile(aes(x=variable,y=id,fill=log2(value)),interpolate=FALSE)+theme_bw()+theme(axis.text.x=element_text(angle=90))+facet_wrap(~stim)+scale_fill_continuous("log2 count",na.value="gray",breaks=round(seq(0,6,l=10)))+theme(axis.text.y=element_text(color=c("black","red")[missed+1]))
p2<-ggplot(toplot2)+geom_tile(aes(x=variable,y=id,fill=log2(value)),interpolate=FALSE)+theme_bw()+theme(axis.text.x=element_text(angle=90,hjust=1))+facet_wrap(~stim)+scale_fill_continuous("log2 count",na.value="gray",breaks=round(seq(0,6,l=10)))+theme(axis.text.y=element_text(color=c("black","red")[2]))
grid.arrange(p1,ncol=1)
grid.arrange(p2,ncol=1)

```{r eval=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/revised_Fluidigm_Multivariate_CCR7_GZMK.pdf",width=6,height=4)
grid.arrange(p1,ncol=1)
dev.off()
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/revised_Fluidigm_Marginal_CCR7_GZMK.pdf",width=6,height=4)
grid.arrange(p2,ncol=1)
dev.off()

A<-cbind(melt(data.frame(matrix(prop.table(as.matrix(CC$n.stim),1)-prop.table(as.matrix(CC$n.unstim),1),ncol=4))),flres$z[,2],1:length(flres$z[,2]))
B<-cbind(melt(data.frame(matrix(prop.table(as.matrix(DD$n.stim),1)-prop.table(as.matrix(DD$n.unstim),1),ncol=2))),flsum$z[,2],1:length(flres$z[,2]))
colnames(A)<-c("Component","Effect Size","Pr(response)","ID")
colnames(B)<-c("Component","Effect Size","Pr(response)","ID")
A$Component<-factor(A$Component,labels=c("-/-","-/+","+/-","+/+"))
B$Component<-factor(B$Component,labels=c("-","+"))
A<-rbind(A,B)
ggplot(A)+geom_point(aes(x=`Effect Size`,y=`Pr(response)`,shape=factor(ID)))+facet_wrap(~Component)+geom_vline(aes(x=0),lty=3)+scale_shape_manual(values=1:16,name="Subject")
```

Simulations
==========
One sided
------
We'll use the model fit from IFNg+ for ENV-1-PTEG based on the ROC and FDR plots
Note that MIMOSA has greater sensitivity and specificity than Fisher's exact test or the Likelihood ratio test. When there are 5 observations, the methods being to perform equivalently.

```{r simulations,cache=FALSE,echo=FALSE,fig.width=5,fig.height=5,include=FALSE}
if(!file.exists("cache/sims.10K.RData")){
sims.50K<-OneSidedSims(hvtn.result.mcmc,N=50000)
sims.25K<-OneSidedSims(hvtn.result.mcmc,N=25000)
sims.10K<-OneSidedSims(hvtn.result.mcmc,N=10000)
cache('sims.50K')
cache('sims.25K')
cache('sims.10K')
}else{
load("cache/sims.10K.RData")
load("cache/sims.25K.RData")
load("cache/sims.50K.RData")
}
ggplot(sims.50K$sim.ROC)+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Average ROC for 10 Data Sets from One-Sided Simulations")+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
ggplot(sims.50K$sim.FDR)+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Observed vs Nominal FDR for 10 Data Sets from One-Sided Simulations")+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
ggplot(sims.25K$sim.ROC)+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Average ROC for 10 Data Sets from One-Sided Simulations")+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
ggplot(sims.25K$sim.FDR)+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Observed vs Nominal FDR for 10 Data Sets from One-Sided Simulations")+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
ggplot(sims.10K$sim.ROC)+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Average ROC for 10 Data Sets from One-Sided Simulations")+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
ggplot(sims.10K$sim.FDR)+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Observed vs Nominal FDR for 10 Data Sets from One-Sided Simulations")+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
```

Two-sided
--------

```{r two.sided.sims, cache=FALSE, echo=FALSE,fig.width=5,fig.height=5,include=FALSE}
if(!file.exists("cache/sims2.10K.RData")){
sims2.50K<-TwoSidedSims(hvtn.result.mcmc,N=50000)
sims2.25K<-TwoSidedSims(hvtn.result.mcmc,N=25000)
sims2.10K<-TwoSidedSims(hvtn.result.mcmc,N=10000)
cache('sims2.50K')
cache('sims2.25K')
cache('sims2.10K')
}else{
load("cache/sims2.10K.RData")
load("cache/sims2.25K.RData")
load("cache/sims2.50K.RData")
}
ggplot(sims2.50K$sim.ROC)+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Average ROC for 10 Data Sets from Two-Sided Simulations")+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
ggplot(sims2.50K$sim.FDR)+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Observed vs Nominal FDR for 10 Data Sets from Two-Sided Simulations")+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
ggplot(sims2.25K$sim.ROC)+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Average ROC for 10 Data Sets from Two-Sided Simulations")+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
ggplot(sims2.25K$sim.FDR)+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Observed vs Nominal FDR for 10 Data Sets from Two-Sided Simulations")+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
smlcnts.roc<-ggplot(subset(sims2.10K$sim.ROC,Nobs==200))+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Average ROC for 10 Data Sets from Two-Sided Simulations (small counts)")+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
smlcnts.fdr<-ggplot(subset(sims2.10K$sim.FDR,Nobs==200))+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Observed vs Nominal FDR for 10 Data Sets from Two-Sided Simulations (small counts)")+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
```

### Figure 3: One-sided and two-sided simulations.

```{r eval=TRUE,echo=FALSE}
levs<-levels(sims.50K$sim.ROC$method)
levs[2]<-c("MIMOSA (one sided)")
sims.50K$sim.ROC$method<-factor(sims.50K$sim.ROC$method,labels=levs)
p1.1<-ggplot(subset(sims.50K$sim.ROC,Nobs==200))+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA (one sided)")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
levs<-levels(sims2.50K$sim.ROC$method)
levs[2]<-c("MIMOSA (two sided)")
sims2.50K$sim.ROC$method<-factor(sims2.50K$sim.ROC$method,labels=levs)
p2.1<-ggplot(subset(sims2.50K$sim.ROC,Nobs==200))+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA (two sided)")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))

levs<-levels(sims.50K$sim.FDR$method)
levs[1]<-"MIMOSA (one sided)"
sims.50K$sim.FDR$method<-factor(sims.50K$sim.FDR$method,labels=levs)
p1.2<-ggplot(subset(sims.50K$sim.FDR,Nobs==200))+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA (one sided)")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
levs<-levels(sims2.50K$sim.FDR$method)
levs[1]<-c("MIMOSA (two sided)")
sims2.50K$sim.FDR$method<-factor(sims2.50K$sim.FDR$method,labels=levs)
p2.2<-ggplot(subset(sims2.50K$sim.FDR,Nobs==200))+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA (two sided)")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)

```{r eval=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_OneSided_ROC_50K.pdf",width=6,height=4)
p1.1
dev.off()
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_TwoSided_ROC_50K.pdf",width=6,height=4)
p2.1
dev.off()
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_TwoSided_FDR_50K.pdf",width=6,height=4)
p2.2
dev.off()
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_OneSided_FDR_50K.pdf",width=6,height=4)
p1.2
dev.off()

```{r echo=FALSE}
p1.1
p2.1
p2.2
p1.2
```

### Multivariate Simulations
-----
Average ROC for 10 multivariate simulations with effect size of 2.5x10-3 and -2.5x10-3 in two of eight components (N=4743 events).

```{r cache=TRUE,results='hide',message=FALSE,warning=FALSE}
set.seed(101)
as<-c(1,1,1,1,3.5,1,1)
au<-c(1,1,1,1,1,3.5,1)
as<-c(1000-sum(as),as)*100
au<-c(1000-sum(au),au)*100
if(!file.exists("mvsims.rds")){
mdsim<-sapply(1:10,function(i){
simd<-MIMOSA:::simMD(alpha.s=as,alpha.u=au,N=100,w=0.5,n=3.5,alternative="not equal")
foo<-.fitMCMC(simd,inits=MIMOSA:::MDMix(simd,initonly=TRUE))
foo
})
saveRDS(mdsim,file="mvsims.rds")
}else{
  mdsim<-readRDS("mvsims.rds")
}
fisher.p<-vector('list',ncol(mdsim))
for(i in 1:ncol(mdsim)){    
  fisher.p[[i]]<-rep(NA,nrow(mdsim[,i]$n.stim))
for(j in 1:nrow(mdsim[,i]$n.stim)){
  fisher.p[[i]][j]<-  fisher.test(cbind(mdsim[,i]$n.stim[j,],mdsim[,i]$n.unstim[j,]),alternative="two.sided",simulate.p.value=TRUE)$p
}
}

ROCfun<-function(x,truth){
  o<-order(x,decreasing=FALSE)
  cbind(tpr=cumsum(truth[o])/sum(truth),fpr=cumsum(!truth[o])/sum(!truth))
}

fisher.p.adj<-lapply(fisher.p,function(x)p.adjust(x,"fdr"))
fisher.roc<-do.call(rbind,by(cbind(do.call(rbind,lapply(fisher.p.adj,function(i)ROCfun(i,gl(2,50)==2)))),rep(gl(100,1),10),function(x)colMeans(x)))
mimosa.adj<-lapply(1:ncol(mdsim),function(i)MIMOSA:::fdr(mdsim[,i]$z))
MIMOSA.roc<-do.call(rbind,by(do.call(rbind,lapply(mimosa.adj,function(i)ROCfun(i,gl(2,50)==2))),rep(gl(100,1),10),function(x)colMeans(x)))
foo<-data.frame(as.data.frame(rbind(fisher.roc,MIMOSA.roc)),Method=gl(2,100,labels=c("Fisher","MIMOSA")))
sz<-element_text(size=18)
p1<-ggplot(foo)+geom_line(aes(x=fpr,y=tpr,col=Method))+scale_y_continuous("TPR")+scale_x_continuous("FPR")+theme_bw()+theme(axis.title.x=sz,axis.title.y=sz,axis.text.x=sz,axis.text.y=sz)
#FDR
mim.fdr<-as.data.frame(cbind(do.call(rbind,lapply(mimosa.adj,function(x)fdrCurves(x,gl(2,50)==2))),rep=gl(100,1)))
mim.fdr<-ddply(mim.fdr,.(rep),summarize,true.fdr.hat=mean(true.fdr),fdr.hat=mean(fdr))
fisher.fdr<-as.data.frame(cbind(do.call(rbind,lapply(fisher.p.adj,function(x)fdrCurves(x,gl(2,50)==2))),rep=gl(100,1)))
fisher.fdr<-ddply(fisher.fdr,.(rep),summarize,true.fdr.hat=mean(true.fdr),fdr.hat=mean(fdr))
foo<-data.frame(rbind(fisher.fdr,mim.fdr),Method=gl(2,100,labels=c("Fisher","MIMOSA")))
p2<-ggplot(foo)+geom_line(aes(y=true.fdr.hat,x=fdr.hat,col=Method))+scale_y_continuous("observed fdr")+scale_x_continuous("nominal fdr")+theme_bw()+theme(axis.title.x=sz,axis.title.y=sz,axis.text.x=sz,axis.text.y=sz)+geom_abline(1,lty=3)
grid.arrange(p1,p2)
```

Competing ROC curves to evaluate the effect of inaccuracies in the true positives and false positives.
-----
Simulate from models with 20,35,50,65,80 percent response rates but treat the ROC evaluation as 100% response.

```{r results='hide',cache=TRUE}
comproc<-CompareROC(hvtn.result.mcmc)
comproc[[1]]$true.resp.rate<-as.numeric(as.character(comproc[[1]]$q))*2
comproc[[2]]$true.resp.rate<-as.numeric(as.character(comproc[[2]]$q))*2

p8<-ggplot(comproc[[1]])+geom_line(aes(x=fdr.hat,y=true.fdr.hat,color=method))+facet_wrap(~true.resp.rate,ncol=5)+theme_bw()+geom_abline(1,lty=3)+labs(title="Observed vs True FDR when \ntrue response rate varies but is assumed to be 100%")+theme(axis.text.x=element_text(angle=90))
p9<-ggplot(comproc[[2]])+geom_line(aes(x=FPR.hat,y=TPR.hat,color=method))+facet_wrap(~true.resp.rate,ncol=5)+theme_bw()+labs(title="ROC curves when \ntrue response rate varies but is assumed to be 100%")+scale_x_continuous("False Positive Rate")+scale_y_continuous("True Positive Rate")+theme(axis.text.x=element_text(angle=90))

library(gridExtra)
```{r eval=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/competeROC.pdf",width=10,height=3)
p8
dev.off()
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/competeROCfdr.pdf",width=10,height=3)
p9
dev.off()

```{r echo=FALSE,fig.width=12,fig.height=6}
p8
p9
```

Supplementary Figure 2
------
Histogram of the empirical proportions of unstimulated cells and overlaid posterior densities of the beta distribution with $\alpha_s$ and $\alpha_u$ estimated from the data for ENV-1-PTEG stimulated, IFN-$\gamma$+ CD4+ T-cells, demonstrating that the assumption of a common distribution for piu across subjects is reasonable.

```{r supp.fig.2,message=FALSE,echo=FALSE}
res<-hvtn.result.mcmc[[7]]
ggplot(data.frame(p.hat=prop.table(as.matrix(res@result@n.unstim),1)[,2]))+geom_histogram(aes(x=p.hat,y=..density..),col="black",fill="gray")+theme_bw()+scale_x_continuous(name=expression(hat(p)[u]))+geom_line(data=data.frame(x=seq(0,3e-4,l=10000),y=dbeta(seq(0,3e-4,l=10000),res@result@params[2,4],res@result@params[2,3]),stim="ENV-1-PTEG"),aes(x=x,y=y),col="blue",lwd=2,lty=3)+facet_wrap(~stim)+theme(strip.text.x=element_text(size=24),axis.text.x=element_text(size=18),axis.text.y=element_text(size=18),axis.title.x=element_text(size=24),axis.title.y=element_text(size=24))
```

Unconstrained MIMOSA fit to data violating model assumptions
-----

```{r violated.assumptions, cache=FALSE,echo=FALSE,fig.width=20,fig.height=10}
sims<-TwoSidedSimsTruncated(hvtn.result.mcmc)
violated.roc<-ggplot(subset(sims$sim.ROC,Num==50000))+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Average ROC for 10 Data Sets from Two-Sided Simulations Violating Model Assumptions")+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Num)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
violated.fdr<-ggplot(subset(sims$sim.FDR,Num==50000))+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+labs(title="Observed vs Nominal FDR for 10 Data Sets from Two-Sided Simulations Violating Model Assumptions")+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Num)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
grid.arrange(violated.roc,violated.fdr,smlcnts.roc,smlcnts.fdr)
```

### Figure S3

```{r eval=TRUE,echo=FALSE}
p1<-ggplot(subset(sims$sim.ROC,Num==50000))+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Num)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
p2<-ggplot(subset(sims$sim.ROC,Num==10000))+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Num)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
p3<-ggplot(subset(sims$sim.FDR,Num==50000))+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Num)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
p4<-ggplot(subset(sims$sim.FDR,Num==10000))+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Num)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)

```{r eval=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_Truncated_ROC_50K.pdf",width=5,height=3)
p1
dev.off()
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_Truncated_ROC_10K.pdf",width=5,height=3)
p2
dev.off()
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_Truncated_FDR_50K.pdf",width=5,height=3)
p3
dev.off()
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_Truncated_FDR_10K.pdf",width=5,height=3)
p4
dev.off()

#rerun one sided sims at 50K, more than 10 reps
sims.50K<-OneSidedSims(hvtn.result.mcmc,N=50000)
levs<-levels(sims.50K$sim.FDR$method)
levs[1]<-"MIMOSA (one sided)"
sims.50K$sim.FDR$method<-factor(sims.50K$sim.FDR$method,labels=levs)
levs<-levels(sims.50K$sim.ROC$method)
levs[2]<-"MIMOSA (one sided)"
sims.50K$sim.ROC$method<-factor(sims.50K$sim.ROC$method,labels=levs)
p5<-ggplot(subset(sims.50K$sim.FDR,Nobs!=200))+geom_line(aes(x=fdr.hat,y=true.fdr.hat,col=relevel(method,"MIMOSA (one sided)")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("Nominal FDR")+scale_y_continuous("Observed FDR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))+geom_abline(1,lty=3)
p6<-ggplot(subset(sims.50K$sim.ROC,Nobs!=200))+geom_line(aes(x=FPR.hat,y=TPR.hat,col=relevel(method,"MIMOSA (one sided)")),direction="vh",lwd=1.5,alpha=0.8)+theme_bw()+scale_x_continuous("FPR")+scale_y_continuous("TPR")+coord_cartesian(xlim=c(-0.01,1.01),ylim=c(-0.01,1.01))+scale_color_discrete("Method")+facet_wrap(~Nobs)+theme(strip.text.x=element_text(size=18),axis.title.x=element_text(size=18),axis.title.y=element_text(size=18),axis.text.x=element_text(angle=90))
```{r eval=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_OneSided_ROC_varyNobs.pdf",width=9,height=3)
p6
dev.off()
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/Sim_OneSided_FDR_varyNobs.pdf",width=8,height=3)
p5
dev.off()
```

### One-sided simulations, 50K events. N=20,50,100

```{r echo=FALSE}
p6
p5
```

### Figure of effect size vs probability of response

```{r }
pd<-do.call(rbind,lapply(hvtn.result.mcmc,function(x)pData(x)))
pd<-cbind(pd,do.call(rbind,lapply(hvtn.result.mcmc,function(x)data.frame(response=MIMOSA:::fdr(x@z)<0.01,Prob.resp=x@z[,2],effect=do.call(c,lapply(x@result@p,function(x)diff(rev(x[,2]))))))))
pd$VACCINE<-factor(pd$RX_CODE%in%c("T1","T2","T3","T4"))
```{r eval=FALSE}
pdf(file="/Users/gfinak/Documents/manuscripts/MIMOSA_Paper/Figures/volcanoplots.pdf",width=15,height=10)
ggplot(subset(pd,TCELLSUB%in%"cd3+/cd4+"))+geom_point(aes(x=effect,y=Prob.resp,color=response,shape=VISITNO:VACCINE))+facet_wrap(TCELLSUB~CYTOKINE,scale="free_x")+theme(axis.text.x=element_text(angle=90))+scale_color_discrete("Response",labels=c("No","Yes"))+theme_bw()
dev.off()
```{r echo=FALSE,fig.width=15}
ggplot(subset(pd,TCELLSUB%in%"cd3+/cd4+"))+geom_point(aes(x=effect,y=Prob.resp,color=response,shape=VISITNO:VACCINE))+facet_wrap(TCELLSUB~CYTOKINE,scale="free_x")+theme(axis.text.x=element_text(angle=90))+scale_color_discrete("Response",labels=c("No","Yes"))+theme_bw()
```

Distribution of total CD4 and CD8 T-cell counts?
-----

```{r echo=FALSE,message=FALSE,warning=FALSE,result='hide',fig.width=5,fig.height=2.5}
suppressPackageStartupMessages(require(scales))
df<-data.frame(tcellsub="CD4",treatment=gl(2,205/2,labels=c("stim","unstim")),count=melt(data.frame(stim=hvtn.result.mcmc[[1]]@result@n.stim[,1],unstim=hvtn.result.mcmc[[1]]@result@n.unstim)[,1]))
df<-rbind(df,data.frame(tcellsub="CD8",treatment=gl(2,224/2,labels=c("stim","unstim")),count=melt(data.frame(stim=hvtn.result.mcmc[[30]]@result@n.stim[,1],unstim=hvtn.result.mcmc[[30]]@result@n.unstim)[,1])))
ggplot(df)+geom_density(aes(x=value,fill=tcellsub),alpha=0.75,position="identity")+theme_bw()+scale_x_continuous(name="Cell Count",labels=comma)
```

Median cell counts for T-cell subsets.
-----

```{r results='asis',message=FALSE,echo=FALSE}
suppressPackageStartupMessages(require(plyr))
suppressPackageStartupMessages(require(reshape))
suppressPackageStartupMessages(require(xtable))
d<-cast(melt(ddply(df,.(tcellsub),summarize,median=round(quantile(value,c(0.5)))),measure="median"),~tcellsub)
print(xtable(d[,-1L],display=c("fg","fg","fg")),type="html",include.rownames=FALSE)
```

Distribution of the proportions for the HVTN data
----

```{r echo=FALSE, message=FALSE, warning=FALSE, result='hide',fig.width=5,fig.height=2.5}
require(scales)
df<-data.frame(tcellsub="CD4",treatment=gl(2,nrow(hvtn.result.mcmc[[1]]@result@n.stim)/2,labels=c("stim","unstim")),melt(data.frame(stim=prop.table(as.matrix(hvtn.result.mcmc[[1]]@result@n.stim),1)[,2],unstim=prop.table(as.matrix(hvtn.result.mcmc[[1]]@result@n.unstim),1)[,2])[,2]))
df<-rbind(df,data.frame(tcellsub="CD8",treatment=gl(2,nrow(hvtn.result.mcmc[[30]]@result@n.unstim)/2,labels=c("stim","unstim")),proportion=melt(data.frame(stim=prop.table(as.matrix(hvtn.result.mcmc[[30]]@result@n.stim),1)[,2],unstim=prop.table(as.matrix(hvtn.result.mcmc[[30]]@result@n.unstim),1)[,2])[,2])))
ggplot(df)+geom_density(aes(x=value,fill=tcellsub),alpha=0.75,position="identity")+theme_bw()+scale_x_continuous(name="Cell Count",labels=scientific)+facet_wrap(~tcellsub,scales="free")
```

Median cell proportions for ICS data.
----

```{r results='asis',message=FALSE,echo=FALSE}
d<-cast(melt(df,measure='value'),~tcellsub,fun.aggregate=function(x)mean(x))[,-1L]
print(xtable(d,display=c("fg","fg","fg")),type="html",include.rownames=FALSE)
```